\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-02-22 analysis/manuscript.Rmd
first draft discussion

With the introduction of central funding mechanisms for publication fees in some European countries since 2012, an increase in hybrid open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
In particular, British (17%), Austrian (15%) and Dutch (13%) universities contributed to this trend. 
However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted into transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
Similar to big deals, transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, but aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023]. -->

@@

By the end of 2023, many transformative agreements were implemented, but interim outcomes are mixed.
The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch]. 
Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023]. 
The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuate market concentration [@Butler_2023; @Shu_2023]. 
Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
The focus on large commercial publishers might also increase inequality, because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest]. 
Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].

The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
For instance, the Association of Swedish Higher Education Institutions (Sveriges universitets- och
högskoleförbund, SUHF) recommended to only support agreements for publishing in full open access journals [^sweden].
Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
In contrast, the German DEAL consortium signed a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028. At the same time, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].

Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this uptake can be attributed to transformative agreements.
Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023]. 
Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.

The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means. 
With this novel and open approach, this the first large-scale analysis aims to answer the following questions:

@@

This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data. 
In the following, these steps are described in more detail.

@@

Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
License information relative to the "accepted manuscript (AM)" version were not considered. 
Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
The biggest differences concerned articles published between 2018 and 2020.
In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs. `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society, Emerald and the Royal Society.
Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023]. After reporting it to OpenAlex, according to the release notes, fixing this issue was still ongoing, which might also explain part of the discrepancy.

@@

## Data Analysis

@@

Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
However, there was also a notable growth in open access provision through other means, most likely publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`). 
<!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->

@@

Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase. 
In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.

@@

Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa as of 2022.
The residual category "Other" includes the remaining countries. 
From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022. 
On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.  
Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020. 
In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022. 
Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals. 
While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.

@@

To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles). 
In contrast, West and Nord European countries published a considerable high number of open access articles, mainly due to transformative agreements. 
Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent. Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100`% in 2022 was comparable to that of the Netherlands, which was `r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%.

Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period. 
Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years. 
In 2022, Sweden had the highest proportion of open access relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020). Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.

>>>>

With the introduction of central funding mechanisms for publication fees in some European countries since 2012, an increase in hybrid open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
In particular, British (17%), Austrian (15%) and Dutch (13%) universities stand out. 
However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted into transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
Similar to big deals, transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, but aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023]. -->

@@

By the end of 2023, many transformative agreements were implemented, but interim outcomes are mixed.
The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch]. 
Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023]. 
The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuate market concentration [@Butler_2023; @Shu_2023]. 
Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
The focus on large commercial publishers might also increase inequality [@Klebel_2023], because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest]. 
Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].

The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
For instance, the Association of Swedish Higher Education Institutions (Sveriges universitets- och
högskoleförbund, SUHF) recommended to only support agreements for publishing in full open access journals [^sweden].
Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
In contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028. At the same time, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].

Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this uptake can be attributed to transformative agreements.
Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023]. 
Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements. 
Likewise, article-level open access invoicing data is only partly available [@Jahn_2021].

The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means. 
With this novel and open approach, this first large-scale analysis aims to answer the following questions:

@@

This study combines data from multiple publicly available data sources as shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data. 
In the following, these steps are described in more detail.

@@

Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
License information relative to the "accepted manuscript (AM)" version were not considered. 
Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].

Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
The biggest differences concerned articles published between 2018 and 2020.
In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs. `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society, Emerald and the Royal Society.
Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023]. After reporting it to OpenAlex, fixing this issue was still ongoing according to the release notes, which might also explain part of the discrepancy.

@@

## Data analysis

@@

Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
However, there was also a notable growth in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`). 
<!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->

@@

Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open access enabled by transformative agreements. 
Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase. 
In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.

@@

Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa as of 2022.
The residual category "Other" includes the remaining countries. 
From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022. 
On the other hand, BRICS recorded a low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.  
Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020. 
In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022. 
Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals. 
While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.

@@

To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles). 
In contrast, West and Nord European countries published a considerable high number of open access articles. 
Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent. Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100`% in 2022 was comparable to that of the Netherlands, which was `r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%.

Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period. 
Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years. 
In 2022, Sweden had the highest proportion of open access relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020). Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.
